### 2020-08-08 pre-analysis-plan/freire-skarbek-lynching.Rmd
update pap, address comments egap and livia

---
title:  |
        | Pre-Analysis Plan:
        | Casting the First Stone: Understanding Attitudes towards Lynching in Brazil[^thanks]
author:
  - Danilo Freire[^freire]
  - David Skarbek[^skarbek]
date: \today
abstract: "Why do citizens support extrajudicial violence? While endorsement for vigilantism has significantly decreased in advanced democracies, lynchings remain tolerated throughout the developing world. In this pre-analysis plan, we propose three survey experiments to investigate (i) which criminal profiles citizens prefer for extralegal punishment, (ii) how individuals justify mob violence, and (iii) whether information provision reduces support for vigilantism. First, we run a conjoint experiment to identify criminal characteristics that are associated with higher likelihood of lynching. In our second experiment, we show respondents a short article about a real lynching episode and ask them whether they see the mob's motivations as legitimate. Lastly, we test whether providing information about the legal consequences of lynching, human rights guarantees, and the risk of retribution makes respondents less likely to endorse extralegal violence. We will run the experiments in Brazil, a country which has seen a sharp rise in vigilante attacks and currently experiences one lynching attempt per day. Our survey will be conducted online and include 2,000 participants."
abstractspacing: double
keywords: Brazil, crime, extralegal violence, lynching, vigilantism
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: template.latex
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# If you need to install any package while knitting the document
r <- getOption("repos")
r["CRAN"] <- "https://cran.rstudio.com/"
options(repos = r)
if (!require("kableExtra")) {
    install.packages("kableExtra")
}
```

[^thanks]: We thank Rosario Aguilar, Miriam Golden, Umberto Mignozzetti, Irfan Nooruddin, Catarina Roman, and Alex Sacco for their valuable comments.

[^freire]: Postdoctoral Research Associate, The Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), <https://danilofreire.github.io>.

[^skarbek]: The Department of Political Science and the Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, USA, [`david_skarbek@brown.edu`](mailto:david_skarbek@brown.edu), <http://davidskarbek.com>.

\newpage

# Motivation
\label{sec:motivation}

\doublespacing

Why do citizens endorse extralegal punishment?[^definitions] Although mob
violence has decreased in advanced democracies, vigilantism remains widespread
in the developing world. In particular, Latin America has been severely affected
by a recent increase in lynchings. Motivated by growing numbers of drug-related
crime and by police ineffectiveness, citizen violence has surged in the region
[@mallen2014vigilantes]. Brazil provides a telling example. From 2011 to 2015,
the country registered about 2,500 lynching episodes, and in 2015 alone 173
people were killed by angry mobs, an average of one execution every two days
[@zechmeister2014political]. According to José Martins, who has studied
lynchings in Brazil for more than 30 years, those figures are not only the
highest in the country's history, but also the highest in the world
[@pearson2018latam]. Victims are generally accused of petty theft, and many are
beaten to death for stealing mobile phones, cooking utensils, or pairs of
sandals [@barbara2015vigilantes]. Lynching perpetrators are often young men,
but they also include female teenagers, elderly women, and even members of
local police forces [@moura2017linchamentos; @oliveira2016mob].

[^definitions]: Here we consider extralegal violence, vigilantism, lynching,
and related terms as synonyms. We follow @moncada2017varieties [6] and define
vigilantism as "the collective use or threat of extralegal violence in
response to an alleged criminal act". While we focus here on vigilantism that
directly involves physical violence, we make no distinction between lethal and
non-lethal lynching episodes.

Recent research suggests that support for vigilantism in Latin America is
associated with exposure to violence [@garcia2019anger], untrustworthy law
enforcement [@zizumbo2017community], perceived insecurity [@ceobanu2011crime;
@godoy2004justice], and social inequality [@arias2010violent]. However, approval
for extrajudicial violence raises important questions about the relationship
between states and societies in the region [@nivette2016institutional]. First,
while vigilantism provides social order to vulnerable populations, it also
undermines the legitimacy of state institutions [@schuberth2013challenging].
Lynching episodes reinforce the idea that civilians should not wait for legal
proceedings and that social control is better exerted by the people themselves
[@black1983crime]. Second, victim characteristics greatly impact the likelihood
and severity of lynchings. In Brazil, a person is rarely lynched when they
commit a crime against males, and females that commit the same crimes as males
also receive less punishment.[^fapesp] Moreover, there is no significant social
difference between victims and perpetrators, nor black Brazilians are
particularly targeted by lynch mobs [@oliveira2016mob]. These dynamics pose a
puzzle to existing explanations for mob violence support and stand in sharp
contrast with the American experience [@smaangs2016doing; @wood2011lynching].

[^fapesp]: See
<https://revistapesquisa.fapesp.br/en/mob-injustice>. Access: June, 2020.

In this study, we endeavour to answer the following questions: 1) which victim
characteristics increase individual support for extralegal punishment?; 2) how
do individuals justify their approval for vigilantism?; and 3) can information
provision reduce support for mob violence? We design a set of three experiments
to tackle these issues. Our first experiment consists of a conjoint analysis
in which we show respondents different profiles of lynching victims and ask them
who they believe deserve the punishment. In our second experiment, we test how
ineffective policing, slow criminal justice systems, and demand for hasher legal
punishment affect individual propensity to support mob violence. Finally, we
assess whether citizens become less supportive of vigilantism when they are
informed about the legal consequences of lynchings, constitutional human rights,
or the risk of _vendettas_ associated with mob violence. While in this
pre-analysis plan we present the three experiments as described above,
their order will be randomised in our survey to mitigate potential order bias.

# Experiment 01
\label{sub:exp01}

We start our analysis with a choice-based conjoint experimental design. We will
present five pairs of criminal profiles to respondents. Each profile consists
of seven attributes: 1) gender of the crime perpetrator; 2) age of the crime
perpetrator; 3) race of the crime perpetrator; 4) offense; 5) gender of the victim
of the motivating crime; 6) age of the victim of the motivating crime; 7)
lynching perpetrators. The attributes and levels are displayed in table
\ref{tab:categories} below.

\vspace{.3cm}

\begin{table}[htpb]
\small
\begin{center}
\caption{\textbf{Attributes and Levels}}
\label{tab:categories}
\begin{tabular}{l !{\vrule width 1pt}p{9cm}}
\Xhline{2\arrayrulewidth}
\textbf{Attribute} & \multicolumn{1}{c}{\textbf{Levels}} \\
\Xhline{2\arrayrulewidth}
\small
Gender of crime perpetrator & Male; female \\ [4pt]
Age of crime perpetrator & Teenager; adult; elderly \\ [4pt]
Race of crime perpetrator & Black; white; native Brazilian; Asian \\ [4pt]
Offense & Picks the pocket; steals the car; molests; rapes; murders \\ [4pt]
Gender of crime victim & Male; female\\ [4pt]
Age of crime victim & Child; teenager; adult; elderly\\ [4pt]
Lynching perpetrators & Bystanders; neighbours; family of the victim; gangs; police \\
\Xhline{2\arrayrulewidth}
\end{tabular}
\end{center}
\end{table}

\newpage

Respondents have to indicate which profile they prefer for extrajudicial
punishment. They are are also required to rate each profile on a scale from 1 to
7, where 1 indicates they definitely believe lynching was not justified and 7
indicates they definitely believe lynching was justified. We take this measure
as a robustness check for our main experimental estimate.

Respondents will read the following prompt before they start the experiment:

* Lynchings are often used as social punishment in Brazil. Lynchings are
  cases in which three or more people physically attack or execute a suspected
  criminal in public. We are interested in knowing more about how Brazilians see
  these episodes. In the next five questions, please read the description of two
  possible lynching victims in Brazil and indicate in which case you believe the
  punishment is more justified. Even if you are not entirely sure, please select
  one of the cases.[^portuguese]

[^portuguese]: We will conduct the survey experiments in Portuguese. The
translation for the first experiment is as follows: Linchamentos são às vezes
usados como punição social no Brasil. Linchamentos são casos nos quais três ou
mais pessoas agridem fisicamente ou executam em público um suspeito de um
crime. Estamos interessados em saber mais sobre como os brasileiros vêem estes
episódios. Nas próximas cinco questões, por favor, leia a descrição de duas
possíveis vítimas de linchamento no Brasil e indique em quais delas você
acredita que a punição é mais justificada. Mesmo que você não tenha certeza,
por favor, escolha um dos casos.

## Sample
\label{sub:sample}

We invite respondents from all regions of Brazil to participate in our
experiment. Qualtrics will recruit 2,000 adult Brazilian citizens to take part
in our survey experiment.

We will also include 15 questions on demographics that we believe may influence
the results. The questions are: 1) the respondent's age, 2) gender, 3) state of
birth, 4) state they currently live in, 5) ethnicity, 6) level of education, 7)
religion (if any), 8) monthly family income in minimum wages, 9) political
orientation (left to right), 10) support for death penalty, 11) whether they had
been victimised in the last year, 12) whether they believe crime has worsened,
13) whether they believe militias improve law and order, 14) trust in the
police, 15) trust in the judicial system.

## Estimation
\label{sub:estimation}

We will estimate our models with the `cregg` package [@leeper2018cregg] for the
R statistical language [@rstats2019]. We will report marginal means and 
average marginal component effects (AMCEs) in our main analysis.
However, @leeper2018subgroup show that AMCEs can be misleading in subgroup
comparisons as model results are sensitive to the choice of reference categories in
interactions. In contrast, marginal means provide a clear description of
quantities of interest, in our case preferences towards lynching, while allowing
for easy comparisons between groups of respondents. So while we will include
both AMCEs and marginal means estimators in our article, we will conduct our
subgroup analysis using only the latter. We will cluster standard errors by
respondent.

## Hypotheses
\label{sec:hypotheses}

We expect that, on average, respondents are more likely to select profiles with
the following attributes levels:

1) Gender of crime perpetrator: male 
2) Age of crime perpetrator: teenager
3) Race of crime perpetrator: blacks
4) Offense: murder
5) Gender of victim: female 
6) Age of victim: child
7) Lynching perpetrators: the family of the victim

We also expect that support for extralegal punishment will be lower for females
of any age, whites and Asians, pick pocketers, and if the victim of the
motivating crime is male. Additionally, we expect that respondents see lynchings
as less legitimate if they are carried out by bystanders, gangs, or the police
forces. We have no prior beliefs about the size of each coefficient.

## Heterogeneous Effects
\label{sub:het01}

We also intend to run subgroup analysis in our sample. We believe that
preferences towards lynching may vary according to age, gender, race, level of
education, political ideology, and previous victimisation. In particular, we
hypothesise that:

* Younger respondents believe that lynchings by the police are less justified.

* Female respondents believe that lynchings against females are less justified.

* Respondents of a particular race are less likely to choose individuals of
  the same race as lynching victims.

* More educated individuals show less support for lynchings against petty
  criminals.

* Political ideology has no effect on support for lynchings when the victim is
  female.

* Right-wingers show stronger support for lynchings against males than
  left-wingers.

* Respondents who were victimised in the last year are more likely to support
  lynchings by neighbours and the police.

# Experiment 02

In our second experiment, we analyse how respondents justify their preferences
for extralegal violence. We assess the impact of three factors that have been
cited as major drivers of vigilante justice in Brazil: 1) police
ineffectiveness; 2) slow criminal justice; 3) demand for harsher punishment for
criminals. Below, we discuss them in further detail.

Research shows that police ineffectiveness frequently appears as a strong
predictor for vigilantism [@cruz2019determinants; @garcia2019anger]. The direct
result of the weakness of police institutions is that citizens decide to take
criminal matters "in their own hands", thus persecuting and punishing the
criminals by themselves. A recent statistic indicates that the police
solves only 10% of the homicides in Brazil, what lends evidence to the link
between weak law enforcement and lynchings [@pearson2018latam].

Another determinant of support for vigilante justice that is suggested in the
literature is the lack of trust in the justice system [@godoy2004justice]. This
is often due to long criminal proceedings, which cause significant anxiety for
the victims. In Brazil, the penal code allows the accused to appeal each
decision several times, so it can take decades before a criminal case is closed
[@sousa2005utilizaccao]. In this respect, citizens do not believe that, even if
the criminal is put to trial, he/she will be punished in a timely matter. Note
that although the police is part of the criminal justice system, we separate the
two institutions in our experiment.

Lastly, we evaluate whether respondents think that the legal punishment
assigned to criminals is not proportional to the severity of their crimes. In
particular, we hope to gauge the demand for iron-fisted criminal justice in
Brazil. Although this treatment arm is related to the previous ones, it
addresses not the efficiency of the institutions, but their legitimacy
[@nivette2016institutional]. We believe that the demand for harsher laws is on
the rise in Brazil, and this may be a major reason why citizens justify their
support for mob violence. In fact, Brazilians are often vocal about their
preference for repressive legal punishment. In a recent article in
_The Wall Street Journal_, a bar owner justified the lynching of the local thug
who killed his son by saying that "_even if he had been put behind bars for 100
years it wouldn't have been enough to pay for all his crimes_"
[@pearson2018latam]. We hypothesise that many Brazilians also share this view.

The experiment consists of three treatment conditions and one control group.
Respondents will read an excerpt of a news article describing a real
lynching case. We have slightly edited the original text so that respondents
have no prior knowledge of the crime.[^jornal] The vignette for the control
group includes no information about the reasons behind the lynching. We will ask
respondents to show their level of support for mob violence using a 0-100
slider, where 0 means no support and 100 means full support. Respondents in each
of the three treatment arms will read the same piece, but with one additional
sentence explaining the motivations behind the lynching. The vignettes are as
follows:

[^jornal]: The original article is available at the following address:
<https://jr.jor.br/2020/05/01/homem-e-linchado-na-vila-progresso>.
Access: August 2020.

* *Control group*: A man was lynched last Friday in Jundiaí, São Paulo.
  According to the neighbours, he tried to break into a house but was
  immobilised and beaten by members of the community.[^control-exp02]

[^control-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade.

* *Treatment 01 - Police ineffectiveness*: A man was lynched last Friday in
  Jundiaí, São Paulo. According to the neighbours, he tried to break into a
  house but was immobilised and beaten by members of the community. **One of
  the residents who took part in the lynching said they had beaten the suspect
  because "the police never patrols the area"**.[^treat01-exp02]

[^treat01-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a polícia
nunca patrulha o local".**

* *Treatment 02 - Criminal justice ineffectiveness*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial system is too slow"**.[^treat02-exp02]

[^treat02-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a justiça
é muito lenta"**.

* *Treatment 03 - Demand for harsher legal punishment*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community. **One
  of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial punishment is not harsh
  enough"**.[^treat03-exp02]

[^treat03-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a punição
da justiça não é dura o suficiente"**.

Before each vignette, respondents will read the following text:

* You will be shown a news article. Please read it carefully. After you read
  the article, we will ask one question about it.[^vig01-exp02]

[^vig01-exp02]: In Portuguese: Uma notícia será apresentada para você. Por favor,
leia a notícia com atenção. Após você ler o artigo, faremos uma pergunta sobre
ele.

After the vignette, respondents will be presented with this question:

* Would you think that the lynching was justified? Please use the slider below
  to indicate your opinion. For disagreement, use 0 to 49; for agreement,
  use 51 to 100. Please use 50 if you neither agree nor
  disagree.[^vig02-exp02]

[^vig02-exp02]: In Portuguese: Você acha que o linchamento foi correto? Por
favor, use a barra abaixo para indicar sua opinião. Para discordar, use de 0 a
49; para concordar, use de 51 a 100. Por favor, use 50 para não concordar nem
discordar.

## Sample

We will randomise the treatment and control conditions to the entire respondent
pool. The randomisation procedure here is independent of that of the previous
experiment.

## Estimation

We will carry out our hypothesis tests with OLS. We will compare the average
score given by the control group with the average score attributed by
respondents in each treatment condition. We will use robust standard errors for
all the models.

## Hypotheses

On average, we expect that:

\vspace{.5cm}

\noindent _H1_: Respondents assigned to the _police ineffectiveness_ treatment
group will show stronger support for lynchings when compared to the control
group. We believe that the effect will be smaller than for the other two
treatment conditions.

\vspace{.3cm}

\noindent _H2_: Respondents assigned to the _criminal justice ineffectiveness_
treatment group will show stronger support for lynchings when compared to the
control group. We hypothesise that the effect for this group will be higher
than that of the first treatment condition but lower than that of the third
treatment group.

\vspace{.3cm}

\noindent _H3_: Respondents assigned to the _demand for harsher punishment_
treatment group will show the strongest support for lynchings.

## Heterogeneous Effects

We hypothesise that these respondents are more likely to agree that lynchings
are justified:

1) Older citizens
2) White
3) Low levels of education
4) Right-wingers
5) Individuals who had been previously victimised

# Experiment 03

Our last experiment aims to measure the effect of information provision on
attitudes about lynching. More specifically, we will test whether reminding
respondents about the legal and social consequences of vigilante justice reduces
the subjects' level of support for such practice.

The experiment has three treatment conditions and a control group. In all of
them we will present respondents with a short statement affirming that some
Brazilians support vigilantism under certain conditions. The control group will
be asked to rate their agreement with the statement with a 0-100 slider where
100 indicates full agreement with the statement and 0 indicates no agreement.
Respondents will be asked to use 0 to 49 if they disagree, 50 if they neither
agree nor disagree, and 50-100 if they agree with the sentence.

Each of the three treatment groups will receive a different message about the
legal or social consequences of lynching in Brazil. In the first treatment
arm, we will inform subjects about how the Brazilian constitution and penal code
punishes civilian violence. The second treatment group will be notified about
the human rights guarantees enshrined in Brazil's legal framework. The last
group will read a vignette that mentions how lynchings can spark _vendettas_ and
initiate a cycle of violence in the community. Subjects in the control group
will receive no information about the consequences of lynchings. The text shown
to the control and treatment groups can be read below.

* *Control group*: In Brazil, some people believe that lynching may be
  justified under certain conditions. To what degree do you agree or disagree
  that lynching can be justified? Please use the slider below to indicate your
  preference. For disagreement, use 0 to 49; for agreement, use 51 to 100.
  Please use 50 if you neither agree nor disagree.[^control-exp03]

[^control-exp03]: In Portuguese: No Brasil, algumas pessoas acreditam que
linchamentos são justificados sob certas condições. O quanto você concorda ou
discorda que linchamentos podem ser justificados? Por favor, use a barra abaixo
para indicar sua preferência. Para indicar que discorda, use de 0 a 49; para
concordar, use de 51 a 100. Por favor, use 50 caso você não concorde nem
discorde.

* *Treatment 01 - Legal punishment for lynching perpetrators*: In Brazil,
  some people believe that lynching may be justified under certain conditions.
  **However, the Brazilian constitution and penal code strictly forbid
  lynching and those involved can be accused of torture or murder**. To what
  degree do you agree or disagree that lynching can be justified? Please use
  the slider below to indicate your preference. For disagreement, use 0 to 49;
  for agreement, use 51 to 100. Please use 50 if you neither agree nor
  disagree.[^treat01-exp03]

[^treat01-exp03]: In Portuguese: No Brasil, algumas pessoas acreditam que
linchamentos são justificados sob certas condições. **Entretanto, a
constituição e o código penal do Brasil proíbem estritamente os linchamentos e
os envolvidos podem ser acusados de tortura ou assassinato.** O quanto você
concorda ou discorda que linchamentos podem ser justificados? Por favor, use a
barra abaixo para indicar sua preferência. Para indicar que discorda, use de 0
a 49; para concordar, use de 51 a 100. Por favor, use 50 caso você não concorde
nem discorde.

* *Treatment 02 - Human rights*: In Brazil, some people believe that lynching
  may be justified under certain conditions. **However, the Brazilian
  constitution states that all individuals have the right of not being
  tortured, including criminals**. To what degree do you agree or disagree that
  lynching can be justified? Please use the slider below to indicate your
  preference. For disagreement, use 0 to 49; for agreement, use 51 to 100.
  Please use 50 if you neither agree nor disagree.[^treat02-exp03]

[^treat02-exp03]: In Portuguese: No Brasil, algumas pessoas acreditam que
linchamentos são justificados sob certas condições. **Entretanto, a
constituição do Brasil afirma que todos os indivíduos têm o direito de não
serem torturados, inclusive criminosos**. O quanto você concorda ou discorda que
linchamentos podem ser justificados? Por favor, use a barra abaixo para indicar
sua preferência. Para indicar que discorda, use de 0 a 49; para concordar, use
de 51 a 100. Por favor, use 50 caso você não concorde nem discorde.

* *Treatment 03 - Vendettas*: In Brazil, some people believe that lynching
  may be justified under certain conditions. **However, lynchings can trigger
  a new cycle of violence as the family or friends of the victim may retaliate
  the community**. To what degree do you agree or disagree that lynching can be
  justified? Please use the slider below to indicate your preference. For
  disagreement, use 0 to 49; for agreement, use 51 to 100. Please use 50 if you
  neither agree nor disagree.[^treat03-exp03]

[^treat03-exp03]: In Portuguese: No Brasil, algumas pessoas acreditam que
linchamentos são justificados sob certas condições. **Entretanto, linchamentos
podem iniciar um ciclo de violência pois a família ou amigos da vítima podem
retaliar a comunidade**. O quanto você concorda ou discorda que linchamentos
podem ser justificados? Por favor, use a barra abaixo para indicar sua
preferência. Para indicar que discorda, use de 0 a 49; para concordar, use de
51 a 100. Por favor, use 50 caso você não concorde nem discorde.

## Sample

We will randomise the treatment and control conditions to all subjects included
in our sample. The treatment assignment here is independent of that of the
previous two experiments.

## Estimation

We are interested in the difference between the average score assigned by each
of the treatment groups minus the average score assigned by the control group.
We will estimate average treatment effects using OLS with dummy indicators for
the treatment groups.

## Hypotheses

In this experiment, we test the following hypotheses:

\vspace{.5cm}

\noindent _H1_: Individuals who receive information about the criminal
consequences of lynching show lower support for vigilantism that those in the
control group.

\vspace{.3cm}

\noindent _H2_: Reminding respondents that torture violates human rights has no
impact on their support for vigilantism.

\vspace{.3cm}

\noindent _H3_: Information about _vendettas_ triggered by lynchings makes
individuals less likely to support vigilantism in comparison with respondents in
the control group. The effect of that information is larger than the other
two treatments.

\vspace{.5cm}

These three hypotheses allow us to test how respondents react to different
consequences of extralegal violence. Hopefully, the experiment may also help
policy-makers design more effective campaigns against vigilantism in Brazil.

## Heterogeneous Effects

As with the previous two experiments, we will measure whether our treatments
have heterogeneous effects. The same demographic characteristics we describe
above -- age, gender, race, education level, political ideology, and previous
victimisation -- will also be used here.

\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}