### 2021-02-08 appendix/appendix.Rmd
add text analysis exp01

---
title: |
       | Supplementary Materials for "Casting the First Stone: Understanding Attitudes towards Lynching in Brazil"
author:
- "Danilo Freire[^freire]"
- "David Skarbek[^skarbek]"
date: \today
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: ../article/references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    toc_depth: 3
    keep_tex: no
    template: template.latex
---

[^freire]: Independent Researcher, <danilofreire@gmail.com>, <http://danilofreire.github.io>. Corresponding author.

[^skarbek]: The Department of Political Science and the Political Theory Project, Brown University, Providence, RI, USA, <davidskarbek@brown.edu>, <http://davidskarbek.com>.

\newpage

# Introduction

This appendix contains the `R` code required to replicate the results we
present in "_Casting the First Stone: Understanding Attitudes towards Lynching
in Brazil_". This file also includes the descriptive statistics of our sample,
the average marginal component effects (AMCEs) for our conjoint experiment,
additional subgroup analyses for all three experiments, and a standard report
which follows APSA's guidelines for experimental research [@gerber2014reporting].

The code below loads the required datasets and the `R` packages we use in our
statistical analyses. It also translates the names of the factor variables from
Portuguese into English.

\footnotesize
```{r loading, warning=F, message=F, error=F, cache=T}
# Install and load required packages
packages <- c("cjoint", "cregg", "estimatr", "grf", "kableExtra", "janitor",
              "quanteda", "seededlda", "stargazer", "tidyverse")
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}
invisible(lapply(packages, library, character.only = TRUE))

# Load the dataset, remove unused rows and columns,
# and convert variable names to snake case
df <- read_csv("../data/data.csv") %>%
  clean_names() %>%
  mutate(response_id            = as.character(response_id),
         consent                = as.factor(q1),
         progress               = as.numeric(progress),
         finished               = as.factor(finished),
         age                    = as.numeric(q2),
         gender                 = as.factor(q3),
         race                   = as.factor(q4),
         education              = as.factor(q5),
         region                 = as.factor(q6),
         household_income       = as.factor(q7),
         ideology               = as.factor(q8),
         death_penalty          = as.factor(q9),
         previous_victim        = as.character(q10),
         previous_victim_text   = as.character(q10_text),
         views_police           = as.factor(q11),
         views_justice          = as.factor(q12),
         exp02_control          = as.numeric(q18),
         exp02_police           = as.numeric(q19),
         exp02_slow_justice     = as.numeric(q20),
         exp02_small_punishment = as.numeric(q21),
         exp03_control          = as.numeric(q22),
         exp03_constitution     = as.numeric(q23),
         exp03_rights           = as.numeric(q24),
         exp03_vendetta         = as.numeric(q25)) %>%
  slice(-1L) %>%
  select(-c(q1:q12, q18:q25)) %>%
  relocate(response_id, consent, progress, finished,
           location_latitude, location_longitude) %>%
  mutate(across(where(is.character), tolower)) %>%
  mutate(across(where(is.factor), tolower))

# Translate factor values from Portuguese to English
df <- df %>%
  mutate(consent = recode(consent,
                          concordo       = "Agree",
                          `não concordo` = "Disagree"),
         gender = recode(gender,
                         "feminino"              = "Female",
                         "masculino"             = "Male",
                         "outro"                 = "Other",
                         "prefiro não responder" = "Rather Not Say"),
         race = recode(race,
                       "amarela"               = "Asian",
                       "branca"                = "White",
                       "indígena"              = "Indigenous",
                       "outra"                 = "Other",
                       "parda"                 = "Mixed Race",
                       "prefiro não responder" = "Rather Not Say",
                       "preta"                 = "Black"),
         race = fct_relevel(race, "Other", "Rather Not Say", after = Inf),
         education = recode(education,
                            "da 1ª à 4ª série do ensino fundamental (antigo primário)" = "Primary School",
                            "da 5ª à 8ª série do ensino fundamental (antigo ginásio)"  = "Secondary School",
                            "ensino médio (antigo 2º grau)"                            = "High School",
                            "ensino superior"                                          = "College",
                            "mestrado ou doutorado"                                    = "Graduate School",
                            "não sei"                                                  = "Don't Know"),
         education = fct_relevel(education, "Primary School", "Secondary School", "High School",
                                 "College", "Graduate School", "Don't Know"),
         region = recode(region,
                         "centro-oeste" = "Center-West",
                         "nordeste"     = "Northeast",
                         "norte"        = "North",
                         "sudeste"      = "Southeast",
                         "sul"          = "South"),
         household_income = recode(household_income,
                                   "acima de r$ 20.000"       = "Above R$20,000",
                                   "até r$ 1.000"             = "Up to R$1,000",
                                   "de r$ 1.001 a r$ 2.000"   = "From R$1,001 to R$2,000",
                                   "de r$ 10.000 a r$ 20.000" = "From R$10,001 to R$20,000",
                                   "de r$ 2.001 a r$ 3.000"   = "From R$2,001 to R$3,000",
                                   "de r$ 3.001 a r$ 5.000"   = "From R$3,001 to R$5,000",
                                   "de r$ 5.001 a r$ 10.000"  = "From R$5,001 to R$10,000"),
         household_income = fct_relevel(household_income, "Up to R$1,000",
                                        "From R$1,001 to R$2,000", "From R$2,001 to R$3,000",
                                        "From R$3,001 to R$5,000", "From R$5,001 to R$10,000",
                                        "From R$10,001 to R$20,000", "Above R$20,000"),
         ideology = recode(ideology,
                           "centro"                = "Center",
                           "centro-direita"        = "Center-Right",
                           "centro-esquerda"       = "Center-Left",
                           "direita"               = "Right",
                           "esquerda"              = "Left",
                           "não sei"               = "Don't Know",
                           "prefiro não responder" = "Rather Not Say"),
         ideology = fct_relevel(ideology, "Left", "Center-Left", "Center",
                                "Center-Right", "Right", "Don't Know",
                                "Rather Not Say"),
         death_penalty = recode(death_penalty,
                                "não"                   = "No",
                                "não sei"               = "Don't Know",
                                "prefiro não responder" = "Rather Not Say",
                                "sim"                   = "Yes"),
         death_penalty = fct_relevel(death_penalty, "Don't Know",
                                     "Rather Not Say", after = Inf),
         views_police = recode(views_police,
                               "boa"                   = "Good",
                               "muito boa"             = "Very Good",
                               "muito ruim"            = "Very Bad",
                               "não sei"               = "Don't Know",
                               "prefiro não responder" = "Rather Not Say",
                               "regular"               = "Regular",
                               "ruim"                  = "Bad"),
         views_police = fct_relevel(views_police, "Very Good", "Good", "Regular",
                                    "Bad", "Very Bad", "Don't Know", "Rather Not Say"),
         views_justice = recode(views_justice,
                                "boa"                   = "Good",
                                "muito boa"             = "Very Good",
                                "muito ruim"            = "Very Bad",
                                "não sei"               = "Don't Know",
                                "prefiro não responder" = "Rather Not Say",
                                "regular"               = "Regular",
                                "ruim"                  = "Bad"),
         views_justice = fct_relevel(views_justice, "Very Good", "Good", "Regular",
                                    "Bad", "Very Bad", "Don't Know", "Rather Not Say"),
         previous_victim_dummy = recode(previous_victim,
                                        "nenhum" = "No",
                                        .missing = NA_character_,
                                        .default = "Yes")) %>%
  relocate(response_id:previous_victim, previous_victim_dummy,
           previous_victim_text:f_5_2_8)

# Check for duplicated values
count(get_dupes(df))
```
\normalsize

# Descriptive Statistics

```{r sample-size, echo=F, warning=F, error=F, message=F, cache=T}
df1 <- df %>% filter(consent == "Agree")
```

We ran our survey experiments from October 30 to December 14 via Qualtrics. Our
sample includes `r nrow(df1)` Brazilians older than 18 years of age from the
five regions of the country (Center-West, North, Northeast, South, and
Southeast). We used quotas for gender and region to ensure that our sample was
similar to the Brazilian population in those characteristics. We also collected
information about whether the subjects had been victimized in the previous 12
months, as well as their opinion the Brazilian judicial system and the police
forces. They follow in the graphs and tables below.

## Informed Consent

About 98% of the interviewees agreed to participate in the survey experiment.
We therefore exclude the remaining cases from the dataset.

\footnotesize
```{r consent, warning=F, error=F, cache=T}
df %>%
  group_by(consent) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 2)) %>%
  rename(Consent = consent) %>%
  kbl(., booktabs = TRUE, caption = "Informed Consent") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")

# Remove subjects who did not agree with consent form
df1 <- df %>% filter(consent == "Agree")
```
\normalsize

## Gender

The gender distribution of our sample is described below. It closely matches
the official data from the [Brazilian Census
Bureau](https://educa.ibge.gov.br/jovens/conheca-o-brasil/populacao/18320-quantidade-de-homens-e-mulheres.html),
which states that women are 51.8% of the population and men comprise 48.2%.

\footnotesize
```{r gender, warning=F, error=F, cache=T}
df1 %>%
  group_by(gender) %>%
  filter(!is.na(gender)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(Gender = gender) %>%
  kbl(., booktabs = TRUE, caption = "Gender") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```
\normalsize

## Age

The age distribution of our sample is shown below. The median age of the survey
respondents is `r round(median(df1$age, na.rm=T), 1)` years old, which
indicates that our sample is older than the Brazilian population (median age $=
33.4$ years old) [@ciafactbook2020].

\footnotesize
```{r age, warning=F, error=F, cache=T, fig.width=6, fig.height=4.3}
tibble(` `    = "Age",
       Median = round(median(df1$age, na.rm = TRUE), 2),
       Mean   = round(mean(df1$age, na.rm = TRUE), 2),
       SD     = round(sd(df1$age, na.rm = TRUE), 2),
       Min    = min(df1$age, na.rm = TRUE),
       Max    = max(df1$age, na.rm = TRUE),
       `NA`   = sum(is.na(df1$age))) %>%
kbl(., booktabs = TRUE, caption = "Age") %>%
row_spec(0, bold = TRUE) %>%
kable_styling(latex_options = "hold_position")

ggplot(subset(df1, !is.na(age)), aes(age)) +
  geom_bar(fill = "#152238") +
  labs(title = "Age Distribution", x = "Age", y = "Count") +
  geom_vline(aes(xintercept = mean(age, na.rm = TRUE)),
            color = "darkred", linetype = 5, size = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```
\normalsize

## Race

The next demographic variable we show here is race. According to the [Brazilian
Census Bureau](https://educa.ibge.gov.br/jovens/conheca-o-brasil/populacao/18319-cor-ou-raca.html),
42.7% of the Brazilian population identify as White, 46.8% as Mixed Race, 9.4%
as Blacks, and 1,1% as Asians or Indigenous. As we see below, our sample
includes more Whites and fewer individuals who identify as Mixed Race. The
number of Blacks roughly coincide with the population statistics.

\footnotesize
```{r race, warning=F, error=F, cache=T}
df1 %>%
  group_by(race) %>%
  filter(!is.na(race)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(Race = race) %>%
  mutate(Race = fct_relevel(Race, "Other", "Rather Not Say", after = Inf)) %>%
  kbl(., booktabs = TRUE,  linesep = "",
      caption = "Race") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```
\normalsize

## Education

```{r educationtext, echo=F, warning=F, error=F, cache=T}
df2 <- df1 %>%
  group_by(education) %>%
  filter(!is.na(education)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(Education = education)
```

As expected, our sample is also more educated than [the Brazilian
population](https://educa.ibge.gov.br/jovens/conheca-o-brasil/populacao/18317-educacao.html).
About `r df2[4, 3]*100`% of the respondents have a college degree, and `r df2[5, 3]`%
have graduate school education.

\footnotesize
```{r education, warning=F, error=F, cache=T}
df1 %>%
  group_by(education) %>%
  filter(!is.na(education)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(Education = education) %>%
  kbl(., booktabs = TRUE,  linesep = "",
      caption = "Education") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```
\normalsize

\newpage

## Household Income

```{r income-text, echo=F, warning=F, error=F, cache=T}
df2 <- df1 %>%
  group_by(household_income) %>%
  filter(!is.na(household_income)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(`Household Income` = household_income)
```

In terms of household income, `r df2[5, 3]*100`% of the respondents earn from
R\$5,0001 to R\$10,000 per month (US\$915 to US\$1830 as of January 2021), which
comprise the largest group in our sample. However, the sample also contains
`r df2[2, 3]*100`% of participants whose household income ranges between R\$1,001
and R\$2,000 (US\$ 184 to US\$368) and `r df2[1, 3]*100`% with household incomes
up to R\$1,000, which is roughly equivalent to Brazil's monthly minimum wage. In
this respect, we have reached participants from all social classes.

\footnotesize
```{r income, warning=F, error=F, cache=T}
df1 %>%
  group_by(household_income) %>%
  filter(!is.na(household_income)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(`Household Income` = household_income) %>%
  kbl(., booktabs = TRUE,  linesep = "",
      caption = "Household Income") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```
\normalsize

## Political Ideology

```{r ideologytext, echo=F, warning=F, error=F, cache=T}
df2 <- df1 %>%
  group_by(ideology) %>%
  filter(!is.na(ideology)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(Ideology = ideology)
```

We have also collected information regarding the subjects' political ideology.
Most respondents identify themselves as right-wingers (`r df2[5, 3]*100`%),
followed by left-wingers (`r df2[1, 3]*100`%), and centrists (`r df2[3, 3]*100`%).
Subjects who do not know their ideology or prefer not to tell their
political beliefs are also large in number (`r df2[6, 3]*100`% and
`r df2[7, 3]*100`%, respectively).

\footnotesize
```{r ideology, warning=F, error=F, cache=T}
df1 %>%
  group_by(ideology) %>%
  filter(!is.na(ideology)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(Ideology = ideology) %>%
  kbl(., booktabs = TRUE,  linesep = "",
      caption = "Political Ideology") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```
\normalsize

## Support for Death Penalty

Below you may find how many respondents support the death penalty.

\footnotesize
```{r death-penalty, warning=F, error=F, cache=T}
df1 %>%
  group_by(death_penalty) %>%
  filter(!is.na(death_penalty)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(`Support for Death Penalty` = death_penalty) %>%
  kbl(., booktabs = TRUE,  linesep = "",
      caption = "Support for Death Penalty") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```
\normalsize

## Previous Victimization

We asked subjects whether they had victimized in the previous 12 months, as
crime victims may be more likely to support lynchings. The responses follow
below.

\footnotesize
```{r victimization, warning=F, error=F, cache=T}
df1 %>%
  group_by(previous_victim_dummy) %>%
  filter(!is.na(previous_victim_dummy)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(Victimization = previous_victim_dummy) %>%
  kbl(., booktabs = TRUE,  linesep = "",
      caption = "Previous Victimization (12 Months)") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```
\normalsize

## Opinion on the Police

Here we show the results for our question on how respondents see the police
forces.

\footnotesize
```{r police, warning=F, error=F, cache=T}
df1 %>%
  group_by(views_police) %>%
  filter(!is.na(views_police)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(`Opinion on the Police` = views_police) %>%
  kbl(., booktabs = TRUE,  linesep = "",
      caption = "Opinion on the Police") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```
\normalsize

## Opinion on the Judicial System

Lastly, we asked how respondents evaluate their local judiciary. As in the
previous question, subjects could choose among five options, as well as affirm
that they do not have an opinion or decline to answer the question.

\footnotesize
```{r justice, warning=F, error=F, cache=T}
df1 %>%
  group_by(views_justice) %>%
  filter(!is.na(views_justice)) %>%
  summarise(N = n()) %>%
  mutate(Frequency = round(N / sum(N), 3)) %>%
  rename(`Opinion on the Judicial System` = views_justice) %>%
  kbl(., booktabs = TRUE,  linesep = "",
      caption = "Opinion on the Judicial System") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```
\normalsize

# Experiment 01

## Description

In our first experiment, we present five pairs of criminal profiles to
respondents. Each profile consists of eight attributes: 1) gender of the crime
perpetrator; 2) age of the crime perpetrator; 3) race of the crime perpetrator;
4) residency of crime perpetrator; 5) offense; 6) gender of the victim of the
motivating crime; 7) age of the victim of the motivating crime; 8) lynching
perpetrators. The attributes and levels are displayed in table
\ref{tab:categories} below.

\vspace{.3cm}

\begin{table}[htpb]
\small
\begin{center}
\caption{\textbf{Attributes and Levels}}
\label{tab:categories}
\begin{tabular}{l !{\vrule width 1pt}p{9cm}}
\Xhline{2\arrayrulewidth}
\textbf{Attribute} & \multicolumn{1}{c}{\textbf{Levels}} \\
\Xhline{2\arrayrulewidth}
\small
Gender of crime perpetrator & Male; female \\ [4pt]
Age of crime perpetrator & Teenager; adult; elderly \\ [4pt]
Race of crime perpetrator & Black; White; Native Brazilian; Asian \\ [4pt]
Residency of crime perpetrator & Resident in the community; outsider \\ [4pt]
Offense & Picks the pocket; steals the car; molests; rapes; murders \\ [4pt]
Gender of crime victim & Male; female\\ [4pt]
Age of crime victim & Child; teenager; adult; elderly\\ [4pt]
Lynching perpetrators & Bystanders; neighbours; family of the victim; gangs; police \\
\Xhline{2\arrayrulewidth}
\end{tabular}
\end{center}
\end{table}

We added three restrictions to the conjoint design to avoid implausible
scenarios. First, female rapists were excluded from the model, but we did
include female molesters in the conjoint experiment. Second, when the offense
was car theft, the victim could not be a child. Lastly, teenagers could not be
victims of car theft either. All other combinations were allowed. We randomized
the attributes using a `.php` script, which is available at
<https://github.com/danilofreire/lynching-experiment-brazil/blob/master/conjoint/portuguese/lynching-conjoint-pt.php>.

Respondents indicated which profile they preferred for extrajudicial
punishment. Prior to the experiments, they had read the following prompt:

* Lynchings are often used as social punishment in Brazil. Lynchings are cases
  in which three or more people physically attack or execute a suspected
  criminal in public. We are interested in knowing more about how Brazilians
  see these episodes. In the next five questions, please read the description
  of two possible lynching victims in Brazil and indicate in which case you
  believe the punishment is more justified. Even if you are not entirely sure,
  please select one of the cases.[^portuguese]

[^portuguese]: Original text in Portuguese: Linchamentos são às vezes usados
como punição social no Brasil. Linchamentos são casos nos quais três ou mais
pessoas agridem fisicamente ou executam em público um suspeito de um crime.
Estamos interessados em saber mais sobre como os brasileiros vêem estes
episódios. Nas próximas cinco questões, por favor, leia a descrição de duas
possíveis vítimas de linchamento no Brasil e indique em quais delas você
acredita que a punição é mais justificada. Mesmo que você não tenha certeza,
por favor, escolha um dos casos.

## Marginal Means Estimator

We estimate the conjoint experiment with the `cregg` package [@leeper2018cregg]
for the `R` statistical language [@rstats2019]. We follow @leeper2020measuring
and report marginal means as our main estimates. Marginal means are easy to
interpret and they are not sensitive to choice of the reference category in
subgroup analyses. The $H_0$ in all models is that the coefficient is equal to
0.5, that is, that respondents are indifferent to that attribute level. Standard
errors are clustered by respondent. The code follows below.

\footnotesize
```{r conjoint-mm, warning=F, error=F, message=F, cache=T, fig.width=7, fig.height=6}
conjoint_data <- read.qualtrics("../data/data-conjoint.csv",
                                responses = c("Q13", "Q14", "Q15",
                                              "Q16", "Q17"),
                                covariates = c("ResponseId",
                                               "Q1", "Q2", "Q3", "Q4",
                                               "Q5", "Q6", "Q7",
                                               "Q8", "Q9", "Q10",
                                               "Q11", "Q12"),
                                new.format = FALSE, respondentID = NULL)

conjoint_data <- conjoint_data %>%
  rename(response_id                      = ResponseId,
         Age                              = Q2,
         Gender                           = Q3,
         Race                             = Q4,
         Education                        = Q5,
         Region                           = Q6,
         "Household Income"               = Q7,
         Ideology                         = Q8,
         "Support death penalty"          = Q9,
         "Previous Victimization"         = Q10,
         "Offense"                        = Crime,
         "Opinion on Policing"            = Q11,
         "Opinion on Judiciary"           = Q12,
         "Gender of crime victim"         = "Gênero.da.vítima",
         "Gender of crime perpetrator"    = "Gênero.do(a).criminoso(a)",
         "Age of crime victim"            = "Idade.da.vítima",
         "Age of crime perpetrator"       = "Idade.do(a).criminoso(a)",
         "Lynching perpetrators"          = "Linchadores",
         "Race of crime perpetrator"      = "Raça.do(a).criminoso(a)",
         "Residency of crime perpetrator" = "Residência.do.criminoso") %>%
  mutate(`Gender of crime perpetrator` = fct_recode(`Gender of crime perpetrator`,
                                                    "Male"   = "Masculino",
                                                    "Female" = "Feminino"),
         `Age of crime perpetrator` = fct_recode(`Age of crime perpetrator`,
                                                 "Teenager" = "Adolescente",
                                                 "Adult"    = "Adulto(a)",
                                                 "Elderly"  = "Idoso(a)"),
         `Race of crime perpetrator` = fct_recode(`Race of crime perpetrator`,
                                                  "Asian"      = "Asiático(a)",
                                                  "White"      = "Branco(a)",
                                                  "Indigenous" = "Indígena",
                                                  "Black"      = "Negro(a)"),
         `Residency of crime perpetrator` = fct_recode(`Residency of crime perpetrator`,
                                                       "Another neighborhood" = "Mora em outro bairro",
                                                       "In the neighborhood"  = "Mora na vizinhança"),
         `Offense` = fct_recode(`Offense`,
                                "Murder"         = "Assassinou",
                                "Pick-pocketing" = "Bateu a carteira",
                                "Rape"           = "Estuprou",
                                "Molestation"    = "Molestou",
                                "Car theft"      = "Roubou o carro"),
         `Gender of crime victim` = fct_recode(`Gender of crime victim`,
                                               " Male"   = "Masculino",
                                               " Female" = "Feminino"),
         `Age of crime victim` = fct_recode(`Age of crime victim`,
                                            " Teenager" = "Adolescente",
                                            " Child"    = "Criança",
                                            " Adult"    = "Adulto(a)",
                                            " Elderly"  = "Idoso(a)"),
         `Lynching perpetrators` = fct_recode(`Lynching perpetrators`,
                                              "Family of the victim" = "Família da vítima",
                                              "Gangs"                = "Gangues",
                                              "Bystanders"           = "Pedestres",
                                              "Police"               = "Polícia",
                                              "Neighbors"            = "Vizinhos")) %>%
  select(-c(16, 18, 20, 22, 24, 26, 28, 30)) %>%
  mutate(response_id = tolower(response_id))

# Model
fm <- selected ~ `Gender of crime perpetrator` +
  `Age of crime perpetrator` + `Race of crime perpetrator` +
  `Residency of crime perpetrator` + `Offense` +
  `Gender of crime victim` + `Age of crime victim` +
  `Lynching perpetrators`

mms <- mm(conjoint_data, fm, id = ~response_id, h0 = 0.5)

# Plot
faces <- c(rep("plain", 5), "bold",
           rep("plain", 4), "bold",
           rep("plain", 2), "bold",
           rep("plain", 5), "bold",
           rep("plain", 2), "bold",
           rep("plain", 4), "bold",
           rep("plain", 3), "bold",
           rep("plain", 2), "bold")

plot(mms, vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "none",
        axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8)

# Table
table_mm <- function(mms, capt) {
dfr <- data.frame(feature = mms[, c(4)],
                  round(mms[, c(5, 6, 8, 9, 10)], digits = 3))
names(dfr) <- c("Feature", "Estimate", "Std. Error",
                "P-Value", "Lower", "Upper")
return(kbl(dfr, "latex", caption = capt, linesep = "",
           booktabs = TRUE) %>%
kable_styling(font_size = 12, full_width = TRUE,
              latex_options = "hold_position") %>%
pack_rows("Gender of crime perpetrator", 1, 2) %>%
pack_rows("Age of crime perpetrator", 3, 5) %>%
pack_rows("Race of crime perpetrator", 6, 9) %>%
pack_rows("Residency of crime perpetrator", 10, 11) %>%
pack_rows("Offense", 12, 16) %>%
pack_rows("Gender of crime victim", 17, 18) %>%
pack_rows("Age of crime victim", 19, 22) %>%
pack_rows("Lynching perpetrators", 23, 27) %>%
column_spec(1, width = "6cm"))
}

table_mm(mms, capt = "Marginal Means -- Full Model")
```
\normalsize

\clearpage

## Average Marginal Component Effect (AMCE) Estimator

We also estimate AMCE coefficients for our conjoint experiment. This method
selects one reference category for each attribute and look at changes from the
baseline level. The reference category are marked as zero in our models.

\footnotesize
```{r conjoint-amce, warning=F, error=F, message=F, cache=T, fig.width=7, fig.height=6}
amces <- cj(conjoint_data, fm, id = ~response_id)

plot(amces, vline = 0.0, header_fmt = "%s") +
  theme(legend.position = "none",
        axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8)

table_mm(amces, capt = "Average Marginal Component Effects -- Full Model")
```
\normalsize

\clearpage

## Subgroup Analyses

In this subsection, we test whether our results vary according to individual
characteristics, such as gender, age, race, income, support for death penalty,
and the respondents' opinions on the judicial system and the police forces. All
models report marginal means. As we shall see, the results are very robust
across all model specifications.

### Gender

Results do not seem to vary according to the gender of the respondent. We focus
here on the differences between males and females and exclude the 11
observations in which respondents preferred not to say their gender or marked
"other" in our questionnaire. Across all conjoint experiment attributes, we see
that an overlap between the 95% confidence intervals for males and females.

\footnotesize
```{r conjoint-females, warning=F, message=F, error=F, cache=T, fig.width=7, fig.height=5.3, fig.pos = "H"}
# Model
cjdt <- full_join(conjoint_data, df1, by = "response_id") %>%
  drop_na(gender) %>%
  filter(gender == c("Male", "Female"))
cjdt$Gender <- factor(cjdt$gender)
mm_by <- cj(cjdt, fm, id = ~response_id, estimate = "mm", h0 = 0.5, by = ~Gender)

# Plot
plot(mm_by, group = "Gender", vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "bottom", axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8)

# Tables
table_mm_by <- function(mm_by, capt) {
dfr <- data.frame(feature = mm_by[, c(5)],
                  round(mm_by[, c(6, 7, 9, 10, 11)], digits = 3))
names(dfr) <- c("Feature", "Estimate", "Std. Error",
                "P-Value", "Lower", "Upper")
return(kbl(dfr, "latex", caption = capt, linesep = "",
           booktabs = TRUE) %>%
kable_styling(font_size = 12, full_width = TRUE,
              latex_options = "hold_position") %>%
pack_rows("Gender of crime perpetrator", 1, 2) %>%
pack_rows("Age of crime perpetrator", 3, 5) %>%
pack_rows("Race of crime perpetrator", 6, 9) %>%
pack_rows("Residency of crime perpetrator", 10, 11) %>%
pack_rows("Offense", 12, 16) %>%
pack_rows("Gender of crime victim", 17, 18) %>%
pack_rows("Age of crime victim", 19, 22) %>%
pack_rows("Lynching perpetrators", 23, 27) %>%
column_spec(1, width = "6cm"))
}

mm_females <- mm_by %>% filter(BY == "Female")
table_mm_by(mm_females, capt = "Marginal Means -- Females")
mm_males <- mm_by %>% filter(BY == "Male")
table_mm_by(mm_males, capt = "Marginal Means -- Males")
```
\normalsize

\clearpage

### Age

As our age variable is continuous, we divide the data into three age brackets:
18-34 years old, 35-54 years old, and 55+ years old. The results show that
seniors (55+) are more likely to select profiles that include murder as an
offense, and less inclined to choose cases involving molestation. The remaining
attributes show little variation.

\footnotesize
```{r conjoint-age, warning=F, message=F, error=F, cache=T, fig.width=7, fig.height=5.2, fig.pos="H"}
# Model
cjdt <- full_join(conjoint_data, df1, by = "response_id") %>%
  drop_na(age) %>%
  mutate(age2 = case_when(age >= 18 & age <= 34 ~ "18-34", age >= 35 & age <= 54 ~ "35-54",
                          age >= 55 ~ "55 plus", TRUE ~ as.character(age)))
cjdt$Age <- factor(cjdt$age2)
mm_by <- cj(cjdt, fm, id = ~response_id, estimate = "mm", h0 = 0.5, by = ~Age)

# Plot
plot(mm_by, group = "Age", vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "bottom", axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8, begin = 0.25)

# Tables
mm_young <- mm_by %>% filter(BY == "18-34")
table_mm_by(mm_young, capt = "Marginal Means -- 18-34 Years Old")
mm_adult <- mm_by %>% filter(BY == "35-54")
table_mm_by(mm_adult, capt = "Marginal Means -- 35-54 Years Old")
mm_senior <- mm_by %>% filter(BY == "55 plus")
table_mm_by(mm_senior, capt = "Marginal Means -- 55+ Years Old")
```
\normalsize

\clearpage

### Race

Below are our results when we disaggregate the data by race. We find that they are
almost identical is all dimensions except for offense. Asian respondents are
much less likely to select profiles that contain pickpocketing as a crime.

\footnotesize
```{r conjoint-race, warning=F, message=F, error=F, cache=T, fig.width=7, fig.height=5.4, fig.pos="H"}
# Model
cjdt <- full_join(conjoint_data, df1, by = "response_id") %>%
  drop_na(race) %>%
  filter(race == c("Asian", "Black", "Mixed Race", "White"))
cjdt$Race <- factor(cjdt$race)
mm_by <- cj(cjdt, fm, id = ~response_id, estimate = "mm", h0 = 0.5, by = ~Race)

# Plot
plot(mm_by, group = "Race", vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "bottom", axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8)

# Tables
mm_asian <- mm_by %>% filter(BY == "Asian")
table_mm_by(mm_asian, capt = "Marginal Means -- Asian")
mm_black <- mm_by %>% filter(BY == "Black")
table_mm_by(mm_black, capt = "Marginal Means -- Black")
mm_mixed <- mm_by %>% filter(BY == "Mixed Race")
table_mm_by(mm_mixed, capt = "Marginal Means -- Mixed Race")
mm_white <- mm_by %>% filter(BY == "White")
table_mm_by(mm_white, capt = "Marginal Means -- White")
```
\normalsize

\clearpage

### Education

Next, we divide our data according to respondents' level of education. As the
number of interviewees with primary or secondary education is low, we merge
them into a single category, while the other levels (high school, college, and
graduate school) remain the same as in our questionnaire.

\footnotesize
```{r conjoint-education, warning=F, message=F, error=F, cache=T, fig.width=8, fig.height=6, fig.pos="H"}
# Model
cjdt <- full_join(conjoint_data, df1, by = "response_id") %>%
  drop_na(education) %>%
  filter(education == c("College", "Graduate School",
                        "Primary School", "Secondary School",
                        "High School")) %>%
  mutate(education2 = case_when(education == "Primary School" ~ "Primary or Secondary School",
                                education == "Secondary School" ~ "Primary or Secondary School",
                                TRUE ~ education),
         education2 = fct_relevel(education2, "Primary or Secondary School",
                                  "High School", "College", "Graduate School"))
cjdt$Education <- factor(cjdt$education2)
mm_by <- cj(cjdt, fm, id = ~response_id, estimate = "mm", h0 = 0.5, by = ~Education)

# Plot
plot(mm_by, group = "Education", vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "bottom", axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8, begin = 0.25)

# Tables
mm_pri_sec <- mm_by %>% filter(BY == "Primary or Secondary School")
table_mm_by(mm_pri_sec, capt = "Marginal Means -- Primary or Secondary School Degree")
mm_high <- mm_by %>% filter(BY == "High School")
table_mm_by(mm_high, capt = "Marginal Means -- High School Degree")
mm_college <- mm_by %>% filter(BY == "College")
table_mm_by(mm_college, capt = "Marginal Means -- College Degree")
mm_grad <- mm_by %>% filter(BY == "Graduate School")
table_mm_by(mm_grad, capt = "Marginal Means -- Graduate School Degree")
```
\normalsize

\clearpage

### Household Income

We also disaggregate the results by monthly household income. As some
categories have few respondents, we group them into three categories: (i) up to
R\$3,000 (US\$550); (ii) from R\$3,001 to R\$5,000 (US\$550-915); and (iii)
above R\$5,000 (US\$915+). The levels roughly represent low, middle, and
high-income households. We find no considerable difference amongst them.

\footnotesize
```{r conjoint-income, warning=F, message=F, error=F, cache=T, fig.width=8, fig.height=6, fig.pos="H"}
# Model
cjdt <- full_join(conjoint_data, df1, by = "response_id") %>%
  drop_na(household_income) %>%
  mutate(household_income2 = case_when(household_income == "Up to R$1,000" ~ "Up to R$3,000",
                                       household_income == "From R$1,001 to R$2,000" ~ "Up to R$3,000",
                                       household_income == "From R$2,001 to R$3,000" ~ "Up to R$3,000",
                                       household_income == "From R$3,001 to R$5,000" ~ "From R$3,001 to R$5,000",
                                       household_income == "From R$5,001 to R$10,000" ~ "Above R$5,000",
                                       household_income == "From R$10,001 to R$20,000" ~ "Above R$5,000",
                                       household_income == "Above R$20,000" ~ "Above R$5,000",
                                       TRUE ~ NA_character_),
         household_income2 = fct_relevel(household_income2, "Up to R$3,000", "From R$3,001 to R$5,000",
                                         "Above R$5,000"))
cjdt$Income <- factor(cjdt$household_income2)
mm_by <- cj(cjdt, fm, id = ~response_id, estimate = "mm", h0 = 0.5, by = ~Income)

# Plot
plot(mm_by, group = "Income", vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "bottom", axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8)

# Tables
mm_3k <- mm_by %>% filter(BY == "Up to R$3,000")
table_mm_by(mm_3k, capt = "Marginal Means -- Up to 3,000 BRL")
mm_5k <- mm_by %>% filter(BY == "From R$3,001 to R$5,000")
table_mm_by(mm_5k, capt = "Marginal Means -- From 3,001 to 5,000 BRL")
mm_abv5k <- mm_by %>% filter(BY == "Above R$5,000")
table_mm_by(mm_abv5k, capt = "Marginal Means -- Above 5,000 BRL")
```
\normalsize

\clearpage

### Support for Death Penalty

Here we assess whether subjects who support the death penalty have different
preferences towards lynching victims. There are fewer respondents who answered
"Don't Know" or "Rather Not Say" to our question, so the confidence intervals
from their estimates are larger than the for the other two categories. The
estimates largely overlap across the four groups, although those who answered
"Rather Not Say" are less favorable to lynching Indigenous criminals.

\footnotesize
```{r conjoint-penalty, warning=F, message=F, error=F, cache=T, fig.width=7, fig.height=5.3, fig.pos="H"}
# Model
cjdt <- full_join(conjoint_data, df1, by = "response_id") %>%
  drop_na(death_penalty) %>%
  mutate(death_penalty = fct_relevel(death_penalty, "Yes", "No",
                                     "Don't Know", "Rather Not Say"))
cjdt$Death_Penalty <- factor(cjdt$death_penalty)
mm_by <- cj(cjdt, fm, id = ~response_id, estimate = "mm", h0 = 0.5, by = ~Death_Penalty)

# Plot
plot(mm_by, group = "Death_Penalty", vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "bottom", axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8)

# Tables
mm_yes <- mm_by %>% filter(BY == "Yes")
table_mm_by(mm_yes, capt = "Marginal Means -- Support for Death Penalty: Yes")
mm_no <- mm_by %>% filter(BY == "No")
table_mm_by(mm_no, capt = "Marginal Means -- Support for Death Penalty: No")
mm_dk <- mm_by %>% filter(BY == "Don't Know")
table_mm_by(mm_dk, capt = "Marginal Means -- Support for Death Penalty: Do Not Know")
mm_rns <- mm_by %>% filter(BY == "Rather Not Say")
table_mm_by(mm_rns, capt = "Marginal Means -- Support for Death Penalty: Rather Not Say")
```
\normalsize

\clearpage

### Previous Victimization

Respondents who had been victimized in the past 12 months also do not have
different preferences towards lynchings victim profiles. The results are
virtually identical for both groups, as one can see below.

\footnotesize
```{r conjoint-victimization, warning=F, message=F, error=F, cache=T, fig.width=7, fig.height=5.6, fig.pos="H"}
# Model
cjdt <- full_join(conjoint_data, df1, by = "response_id") %>%
  mutate(previous_victim_dummy, "Yes", "No")
cjdt$Previous_Victim <- factor(cjdt$previous_victim_dummy)
mm_by <- cj(cjdt, fm, id = ~response_id, estimate = "mm", h0 = 0.5, by = ~Previous_Victim)

# Plot
plot(mm_by, group = "Previous_Victim", vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "bottom", axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8)

# Tables
mm_yes <- mm_by %>% filter(BY == "Yes")
table_mm_by(mm_yes, capt = "Marginal Means -- Previous Victimization (12 Months): Yes")
mm_no <- mm_by %>% filter(BY == "No")
table_mm_by(mm_no, capt = "Marginal Means -- Previous Victimization (12 Months): No")
```
\normalsize

\clearpage

### Opinion on the Police

Experimental results do not change when we break the responses according to how
subjects view the police forces.

\footnotesize
```{r conjoint-policing, warning=F, message=F, error=F, cache=T, fig.width=8, fig.height=6, fig.pos="H"}
# Model
cjdt <- full_join(conjoint_data, df1, by = "response_id") %>%
  mutate(views_police2 = case_when(views_police == "Rather Not Say" ~ NA_character_,
                                   views_police == "Don't Know" ~ NA_character_,
                                   TRUE ~ views_police),
         views_police2 = fct_relevel(views_police2, "Very Good", "Good",
                                     "Regular", "Bad", "Very Bad")) %>%
  drop_na(views_police2)
cjdt$Police <- factor(cjdt$views_police2)
mm_by <- cj(cjdt, fm, id = ~response_id, estimate = "mm", h0 = 0.5, by = ~Police)

# Plot
plot(mm_by, group = "Police", vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "bottom", axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8)

# Tables
mm_vgood <- mm_by %>% filter(BY == "Very Good")
table_mm_by(mm_vgood, capt = "Marginal Means -- Opinion on the Police: Very Good")
mm_good <- mm_by %>% filter(BY == "Good")
table_mm_by(mm_good, capt = "Marginal Means -- Opinion on the Police: Good")
mm_regular <- mm_by %>% filter(BY == "Regular")
table_mm_by(mm_regular, capt = "Marginal Means -- Opinion on the Police: Regular")
mm_bad <- mm_by %>% filter(BY == "Bad")
table_mm_by(mm_bad, capt = "Marginal Means -- Opinion on the Police: Bad")
mm_vbad <- mm_by %>% filter(BY == "Very Bad")
table_mm_by(mm_vbad, capt = "Marginal Means -- Opinion on the Police: Very Bad")
```
\normalsize

\clearpage

### Opinion on the Judicial System

Lastly, we analyze whether personal beliefs about the judicial system affect
the type of lynching victim respondents select.

\footnotesize
```{r conjoint-justice, warning=F, message=F, error=F, cache=T, fig.width=8, fig.height=5.6, fig.pos="H"}
# Model
cjdt <- full_join(conjoint_data, df1, by = "response_id") %>%
  mutate(views_justice2 = case_when(views_justice == "Rather Not Say" ~ NA_character_,
                                    views_justice == "Don't Know" ~ NA_character_,
                                   TRUE ~ views_justice),
         views_justice2 = fct_relevel(views_justice2, "Very Good", "Good",
                                      "Regular", "Bad", "Very Bad")) %>%
  drop_na(views_justice2)
cjdt$Judiciary <- factor(cjdt$views_justice2)
mm_by <- cj(cjdt, fm, id = ~response_id, estimate = "mm", h0 = 0.5, by = ~Judiciary)

# Plot
plot(mm_by, group = "Judiciary", vline = 0.5, header_fmt = "%s") +
  theme(legend.position = "bottom", axis.text.y = element_text(face = faces, size = 10)) +
  scale_colour_viridis_d(option = "inferno", end = 0.8)

# Tables
mm_vgood <- mm_by %>% filter(BY == "Very Good")
table_mm_by(mm_vgood, capt = "Marginal Means -- Opinion on the Judicial System: Very Good")
mm_good <- mm_by %>% filter(BY == "Good")
table_mm_by(mm_good, capt = "Marginal Means -- Opinion on the Judicial System: Good")
mm_regular <- mm_by %>% filter(BY == "Regular")
table_mm_by(mm_regular, capt = "Marginal Means -- Opinion on the Judicial System: Regular")
mm_bad <- mm_by %>% filter(BY == "Bad")
table_mm_by(mm_bad, capt = "Marginal Means -- Opinion on the Judicial System: Bad")
mm_vbad <- mm_by %>% filter(BY == "Very Bad")
table_mm_by(mm_vbad, capt = "Marginal Means -- Opinion on the Judicial System: Very Bad")
```
\normalsize

\clearpage

## Text Analysis

In addition to the conjoint experiments, we also asked respondents to justify
their profile choices. We added a textbook after each conjoint and informed
subjects that their responses were optional. However, we obtained `r (nrow(df1) - sum(is.na(df1$q13_text))) + (nrow(df1) - sum(is.na(df1$q14_text))) + (nrow(df1) - sum(is.na(df1$q15_text))) + (nrow(df1) - sum(is.na(df1$q16_text))) + (nrow(df1) - sum(is.na(df1$q17_text)))` responses
in our survey, which we analyze here.

First, we concatenate all text responses into a single vector. Then we tokenize
the sentences, remove Portuguese stop words and punctuation, and select the
words that appear most frequently in the texts.

The graphs shows that _crime_ (same as in English), _porque_ (because),
_linchamento_ (lynching), and _caso_ (case) are the words respondents use most
often. This is expected as subjects were asked to justify their choices. The
next words in the list are related to victim or crime characteristics, such as
_criança_ (child), _estupro_ (rape), _assassinato_ (murder), and _vítima_
(victim). Indeed, they provide evidence our previous findings and confirm that
respondents select lynching victim profiles according to these two factors.
Criminal characteristics, such as age or race, do not seem to be particularly
relevant, as respondents do not mention them as much. The next words are
_nenhum_ (none), _contra_ (against), _acho_ (I think), _casos_ (cases), _ter_
(have to), and _justiça_ (justice). We believe these words correspond to cases
where respondents wanted to affirm they do not have any preference regarding
the lynching profiles, or that they would rather not have chosen any of the
alternatives.

\footnotesize
```{r conjoint-text, warning=F, message=F, error=F, cache=T, fig.width=8, fig.height=5.6, fig.pos="H", fig.cap = "Words respondents use most frequently when justifying their choices of lynching profiles"}
dfmat <- df1 %>%
  select(q13_text, q14_text, q15_text, q16_text, q17_text) %>%
  gather() %>%
  drop_na() %>%
  corpus(text_field = "value") %>%
  tokens(remove_punct = TRUE, remove_numbers = TRUE,
         remove_symbol = TRUE) %>%
  dfm() %>%
  dfm_remove(., pattern = c(stopwords("pt", source = "snowball"),
                            "é", "ser"))

# Plot
dfmat %>%
  textstat_frequency(n = 15) %>%
  ggplot(aes(x = reorder(feature, frequency), y = frequency)) +
  geom_point(colour = "#152238") +
  coord_flip() +
  labs(x = NULL, y = "Frequency") +
  theme_minimal()
```
\normalsize

We also construct a feature co-occurrence matrix (FCM) that shows which words
appear together in the responses we collected. Again, the results confirm the
findings of the conjoint experiment. As suggested in the previous graph, we see
a central cluster that describe crime and victim characteristics and includes
the words _linchamentos_ (lynchings), _caso_ (case), _estupro_ (rape),
_criança_ (child), _assassinato_ (murder), and _vítima_ (victim). This
highlights that these are the most important reasons why respondents choose
lynching profiles.

We also note that there is another word cluster on the left. It contains words
that indicate that some respondents do not support lynchings, such as _não_
(no), _opção_ (choice), _nada_ (nothing), _justifica_ (justifies),
_justificável_ (justifiable), and _escolher_ (choose).

\footnotesize
```{r conjoint-fcm, warning=F, message=F, error=F, cache=T, fig.width=8, fig.height=5.6, fig.pos="H", fig.cap = "Semantic network analysis"}
fcmat <- fcm(dfmat)
feat <- names(topfeatures(fcmat, 70))
fcmat_select <- fcm_select(fcmat, pattern = feat, selection = "keep")
size <- log(colSums(dfm_select(dfmat, feat, selection = "keep")))

# Plot
set.seed(144)
textplot_network(fcmat_select, min_freq = 0.8, vertex_size = size / max(size) * 3, max.overlaps = 30)
```
\normalsize

Lastly, we estimate a latent Dirichlet allocation (LDA) model to identify the
three most important topics in our corpus. The first topic includes words that
describes lynching perpetrators, as it contains the words like _polícia_
(police), _pessoas_ (people), and _família_ (family). The second topic
identifies the same words we associate with respondents who are against
lynchings, such as _nenhum_ (none), _opção_ (choice), _não_ (no), and _contra_
(against). The last topic, in contrast, refer to victim and crime
characteristics, which have also appeared in our previous estimations. Some of
the most common words in this group are _crime_ (crime), _criança_ (child),
_estupro_ (rape), _vítima_ (victim), _porque_ (because), _idoso_ (elderly),
_grave_ (serious), and _molestar_ (molest).

As our results show, respondents decide which individual deserves punishment
based on factors related to the crime he/she committed, specially the crime
victim. There is also a group of respondents that oppose lynchings in
principle, who affirm that lynchings are not justified.

\footnotesize
```{r conjoint-lda, warning=F, message=F, error=F, cache=T, fig.width=8, fig.height=5.6, fig.pos="H"}
# Latent Dirichlet allocation
tmod_lda <- textmodel_lda(dfmat, k = 3)
terms(tmod_lda, 10)
```
\normalsize

# Experiment 02

## Description

In our second experiment, we analyze how respondents justify their preferences
towards extralegal violence. We assess the impact of three factors that have
been cited as major drivers of lynchings: 1) police ineffectiveness; 2) slow
criminal justice; 3) demand for harsher punishment for criminals. Below, we
discuss them in further detail.

Research shows that police ineffectiveness frequently appears as a strong
predictor of vigilantism [@cruz2019determinants; @garcia2019anger]. The direct
result of the weakness of police institutions is that citizens decide to take
criminal matters "into their own hands", thus persecuting and punishing the
criminals by themselves. A recent statistic indicates that the police solves
only 10% of the homicides in Brazil, which lends support to the link between
weak law enforcement and lynchings [@pearson2018latam].

Another possible determinant of lynching support is lack of trust in the
justice system [@godoy2004justice; @smith2019contradictions]. This is often due
to long criminal proceedings, which cause significant anxiety for the victims.
In Brazil, the penal code allows the accused to appeal each decision several
times, so it can take decades before a criminal case is closed
[@sousa2005utilizaccao]. In this respect, citizens do not believe that, even if
the criminal is put to trial, he/she will be punished in a timely matter. Note
that although the police is technically part of the criminal justice system, we
analyze the two institutions separately in our experiment.

Lastly, we evaluate whether respondents think that the legal punishment
assigned to criminals is not proportional to the severity of their crimes. In
particular, we intend to gauge the demand for iron-fisted criminal justice in
Brazil. Although this treatment arm is related to the previous ones, it
addresses not the efficiency of the institutions, but their legitimacy
[@nivette2016institutional]. In fact, Brazilians are often vocal about their
preference for repressive legal punishment. In a recent article in _The Wall
Street Journal_, a bar owner justified the lynching of the local thug who
killed his son by saying that "_even if he had been put behind bars for 100
years it wouldn't have been enough to pay for all his crimes_"
[@pearson2018latam]. We hypothesize that many Brazilians also share this view.

The experiment consists of three treatment conditions and one control group.
Respondents read an excerpt of a news article describing a real lynching case.
We have slightly edited the original text so that respondents have no prior
knowledge of the crime.[^jornal] The vignette for the control group includes no
information about the reasons behind the lynching. We ask respondents to show
their level of lynching support using a 0-100 slider, where 0 means no support
and 100 means full support. Respondents in each of the three treatment arms
read the same piece, but with one additional sentence explaining the
motivations behind the lynching. The vignettes are as follows:

[^jornal]: The original article is available at the following address:
<https://jr.jor.br/2020/05/01/homem-e-linchado-na-vila-progresso>.
Access: August 2020.

* *Control group*: A man was lynched last Friday in Jundiaí, São Paulo.
  According to the neighbours, he tried to break into a house but was
  immobilised and beaten by members of the community.[^control-exp02]

[^control-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade.

* *Treatment 01 - Police ineffectiveness*: A man was lynched last Friday in
  Jundiaí, São Paulo. According to the neighbours, he tried to break into a
  house but was immobilised and beaten by members of the community. **One of
  the residents who took part in the lynching said they had beaten the suspect
  because "the police never patrols the area"**.[^treat01-exp02]

[^treat01-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a polícia
nunca patrulha o local".**

* *Treatment 02 - Criminal justice ineffectiveness*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial system is too slow and the perpetrator is on
  the street until the case is heard"**.[^treat02-exp02]

[^treat02-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a justiça
é muito lenta e os criminosos ficam soltos até o julgamento"**.

* *Treatment 03 - Demand for harsher legal punishment*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial punishment is not harsh
  enough"**.[^treat03-exp02]

[^treat03-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a punição
da justiça não é dura o suficiente"**.

Before each vignette, respondents will read the following text:

* You will be shown a news article. Please read it carefully. After you read
  the article, we will ask you one question about it.[^vig01-exp02]

[^vig01-exp02]: In Portuguese: Uma notícia será apresentada para você. Por
favor, leia a notícia com atenção. Após você ler o artigo, faremos uma pergunta
sobre ele.

After the vignette, respondents will be presented with this question:

* Would you think that the lynching was justified? Please use the slider below
  to indicate your opinion. For disagreement, use 0 to 49; for agreement,
  use 51 to 100. Please use 50 if you neither agree nor
  disagree.[^vig02-exp02]

[^vig02-exp02]: In Portuguese: Você acha que o linchamento foi correto? Por
favor, use a barra abaixo para indicar sua opinião. Para discordar, use de 0 a
49; para concordar, use de 51 a 100. Por favor, use 50 para não concordar nem
discordar.

## Main Results

Table \ref{tab:exp02main} summarizes our results. As we see, none of the
treatment effects are statistically significant at conventional levels. This
indicates that institutional factors do not affect lynching support in our
sample.

\footnotesize
```{r exp02-main, warning=F, message=F, error=F, cache=T, fig.pos="H", results='asis'}
df_exp02 <- df1 %>%
  mutate(exp02_outcomes = coalesce(exp02_control, exp02_police,
                                   exp02_slow_justice, exp02_small_punishment),
         exp02_any_treat = case_when(!is.na(exp02_control) ~ "0",
                                     !is.na(exp02_police) ~ "1",
                                     !is.na(exp02_slow_justice) ~ "1",
                                     !is.na(exp02_small_punishment) ~ "1",
                                     TRUE ~ NA_character_),
         exp02_police_treat = case_when(!is.na(exp02_control) ~ "0",
                                        !is.na(exp02_police) ~ "1"),
         exp02_slow_justice_treat = case_when(!is.na(exp02_control) ~ "0",
                                              !is.na(exp02_slow_justice) ~ "1"),
         exp02_small_punishment_treat = case_when(!is.na(exp02_control) ~ "0",
                                                  !is.na(exp02_small_punishment) ~ "1"))

m1 <- lm(exp02_outcomes ~ exp02_police_treat, data = df_exp02)
m2 <- lm(exp02_outcomes ~ exp02_slow_justice_treat, data = df_exp02)
m3 <- lm(exp02_outcomes ~ exp02_small_punishment_treat, data = df_exp02)
m4 <- lm(exp02_outcomes ~ exp02_any_treat, data = df_exp02)

stargazer(m1, m2, m3, m4, se = starprep(m1, m2, m3, m4),
          p = starprep(m1, m2, m3, m4, se_type = "HC2", stat = "p.value"),
          header = FALSE, align = TRUE, label = "tab:exp02main",
          title = "Experiment 02 -- Main Results", style = "apsr",
          dep.var.labels = "\\textbf{Lynching Support}\\vspace{.5cm}",
          covariate.labels = c("Police does not patrol area", "Justice too slow",
                               "Punishment not harsh enough", "Combined treatments"),
          column.sep.width = "3pt", notes = "Robust standard errors in parentheses.",
          keep.stat = "n", no.space = TRUE, font.size = "footnotesize")
```
\normalsize

\clearpage

## Demographic Characteristics and Lynching Support

Here we evaluate the impact of individual characteristics on lynching support.
We combine all treatment and control groups and measure whether demographic
information, which we collected in our questionnaire, changes baseline levels.

\footnotesize
```{r exp02-demographics, warning=F, message=F, error=F, cache=T, fig.pos="H", results='asis'}
df_exp02_group <- df_exp02 %>%
  filter(gender %in% c("Female", "Male"),
         race %in% c("Asian", "Black", "Mixed Race", "White", "Indigenous"),
         education %in% c("College", "Graduate School", "Primary School",
                          "Secondary School", "High School")) %>%
  mutate(race = fct_relevel(race, "White", "Black", "Mixed Race", "Asian",
                            "Indigenous")) %>%
  mutate(education2 = case_when(education == "Primary School" ~ "Primary or Secondary School",
                                education == "Secondary School" ~ "Primary or Secondary School",
                                TRUE ~ as.character(education)),
         education2 = fct_relevel(education2, "Primary or Secondary School",
                                  "High School", "College", "Graduate School")) %>%
  mutate(views_police = fct_relevel(views_police, "Regular", "Very Good", "Good",
                                    "Bad", "Very Bad")) %>%
  mutate(views_justice = fct_relevel(views_justice, "Regular", "Very Good", "Good",
                                    "Bad", "Very Bad")) %>%
  mutate(ideology = fct_relevel(ideology, "Center", "Left", "Center-Left",
                                "Center-Right", "Right", "Don't Know",
                                "Rather Not Say")) %>%
  mutate(household_income2 = case_when(household_income == "Up to R$1,000" ~ "Up to R$3,000",
                                       household_income == "From R$1,001 to R$2,000" ~ "Up to R$3,000",
                                       household_income == "From R$2,001 to R$3,000" ~ "Up to R$3,000",
                                       household_income == "From R$3,001 to R$5,000" ~ "From R$3,001 to R$5,000",
                                       household_income == "From R$5,001 to R$10,000" ~ "Above R$5,000",
                                       household_income == "From R$10,001 to R$20,000" ~ "Above R$5,000",
                                       household_income == "Above R$20,000" ~ "Above R$5,000",
                                       TRUE ~ NA_character_),
  household_income2 = fct_relevel(household_income2, "Up to R$3,000", "From R$3,001 to R$5,000",
                                  "Above R$5,000"))

# Models
m1 <- lm(exp02_outcomes ~ gender, data = df_exp02_group)
m2 <- lm(exp02_outcomes ~ age, data = df_exp02_group)
m3 <- lm(exp02_outcomes ~ race, data = df_exp02_group)
m4 <- lm(exp02_outcomes ~ education2, data = df_exp02_group)
m5 <- lm(exp02_outcomes ~ household_income2, data = df_exp02_group)
m6 <- lm(exp02_outcomes ~ ideology, data = df_exp02_group)
m7 <- lm(exp02_outcomes ~ previous_victim_dummy, data = df_exp02_group)
m8 <- lm(exp02_outcomes ~ views_police, data = df_exp02_group)
m9 <- lm(exp02_outcomes ~ views_justice, data = df_exp02_group)

# Tables (robust standard errors from estimatr)
stargazer(m1, m2, m3, m4, m5, se = starprep(m1, m2, m3, m4, m5),
          p = starprep(m1, m2, m3, m4, m5, se_type = "HC1", stat = "p.value"),
          header = FALSE, align = TRUE, label = "tab:exp02demo1",
          title = "Experiment 02 -- Demographic Characteristics and Lynching Support",
          dep.var.labels = "\\textbf{Lynching Support}\\vspace{.5cm}", style = "apsr",
          column.labels = c("Gender", "Age", "Race", "Education", "Income"),
          covariate.labels = c("Male", "Years", "Black", "Mixed Race",
                               "Asian", "Indigenous", "High School", "College",
                               "Graduate School", "3,001 to 5,000 BRL", "5,000+ BRL"),
          column.sep.width = "2pt", notes = c("Reference categories: Model 1: Female;
                                              Model 3: White; Model 4: Primary or secondary school;",
                                              "Model 5: Up to 3,000 BRL. Robust standard errors in
                                              parentheses."),
          keep.stat = "n", no.space = TRUE, font.size = "footnotesize")

```
\clearpage

```{r exp02-demographics2, warning=F, message=F, error=F, cache=T, fig.pos="H", results='asis'}
stargazer(m6, m7, m8, m9, se = starprep(m6, m7, m8, m9),
          p = starprep(m6, m7, m8, m9, se_type = "HC2", stat = "p.value"),
          header = FALSE, align = TRUE, label = "tab:exp02demo2",
          title = "Experiment 02 -- Demographic Characteristics and Lynching Support",
          dep.var.labels = "\\textbf{Lynching Support}\\vspace{.5cm}", style = "apsr",
          column.labels = c("Ideology", "Previous Victimization", "Police",
                            "Judiciary"),
          covariate.labels = c("Left", "Center-Left", "Center-Right", "Right",
                               "Don't Know", "Rather Not Say", "Yes",
                               "Very Good", "Good", "Bad", "Very Bad",
                               "Don't Know", "Rather Not Say",
                               "Very Good", "Good", "Bad", "Very Bad",
                               "Don't Know", "Rather Not Say"),
          column.sep.width = "2pt", notes = c("Reference categories: Model 1: Center;
                                              Model 2: No; Model 3: Regular; Model 4: Regular.",
                                              "Robust standard errors in parentheses."),
          keep.stat = "n", no.space = TRUE, font.size = "footnotesize")
```
\normalsize

## Subgroup Analyses

\footnotesize
```{r exp02-forest, warning=F, error=F, message=F, cache=T, fig.width=7, fig.height=6, eval=F}
df_exp02_police <- df_exp02_group %>%
  select(exp02_outcomes, exp02_police_treat, gender, age, race, education, household_income2, ideology, views_police, views_justice) %>%
  mutate(exp02_police_treat = as.numeric(exp02_police_treat))

  drop_na()

library(FindIt)
F1 <-FindIt(model.treat = exp02_outcomes ~ exp02_police_treat,
            model.main = ~ race,
            model.int = ~ race,
            data = df_exp02_police,
            type = "binary",
            treat.type = "single")

F1 <-FindIt(model.treat= outcome ~ treat,
            model.main= ~ age+educ+black+hisp+white+marr+nodegr+log.re75+u75,
            model.int= ~ age+educ+black+hisp+white+marr+nodegr+log.re75+u75, data = LaLonde, type="binary", treat.type="single")

cases <- sample(seq_len(nrow(df_exp02_police)), round(nrow(df_exp02_police) * .6))
train <- df_exp02_police[cases, ]
test <- df_exp02_police[-cases, ]

x <- c("age")
g <- c("gender", "race", "education", "household_income2",
"ideology", "views_police", "views_justice")
X2 <- model.matrix(~ 0 + ., train[c(x, g)])

cf <- causal_forest(
  X = X2,
  Y = as.numeric(train$exp02_outcomes),
  W = as.integer(train$exp02_police_treat),
  num.trees = 1000,
  seed = 1839
)

cf %>%
  variable_importance() %>%
  as.data.frame() %>%
  mutate(variable = colnames(cf$X.orig)) %>%
  arrange(desc(V1))

preds <- predict(
  object = cf,
  newdata = model.matrix(~ ., data = test[, 3:ncol(train)]),
  estimate.variance = TRUE
)

test$preds <- preds$predictions

ggplot(test, aes(x = views_justice, y = preds)) +
  geom_point() +
  geom_smooth(method = "loess", span = 1) +
  theme_light()

average_treatment_effect(cf, target.sample = "all")
tree <- get_tree(cf, 1)
plot(tree)

dr_scores <- double_robust_scores(cf)
tree <- policy_tree(X2, dr_scores, 3)
plot(tree)
plot_htes(preds, ci = TRUE)
```

\newpage

# Session Information
\label{sec:session-info}

\footnotesize
```{r session, cache=TRUE}
sessionInfo()
```
\normalsize

\clearpage