### 2022-09-08 cover-letter/cover-letter.Rmd
update submission

---
author: 
    - Danilo Freire
opening: Dear Editors of Research & Politics,
closing: Sincerely,
address:
    - \textbf{The Editors}
    - Research & Politics
return-address: 
    - Danilo Freire
    - University of Lincoln
    - Brayford Pool, Lincoln LN6 7TS
    - United Kingdom
    - dfreire@lincoln.ac.uk
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: onehalf
margin: 2cm
papersize: a4paper
fontsize: 12pt
inline-links: true
colorlinks: true
output: 
    pdf_document:
        template: letter-template.latex
---

\doublespacing

I am writing to submit our research note entitled "Vigilantism and
Institutions: Understanding Attitudes toward Lynching in Brazil" for
consideration as a _Research & Politics_ research article. We use two survey
experiments and collect primary quality evidence to examine in which
circumstances individuals perceive lynchings as a justified extrajudicial
punishment and whether providing different types of information about lynching
reduces support for it. Our results show that respondents strongly support
lynchings perpetrated by the family of a victim of sexual abuse or murder. We
also find that that popular support for lynchings can be substantially reduced
by reminding individuals of the penalties for lynching and of the risk of
sparking vendettas, but not by emphasizing human rights protections.

We believe out findings sheds new light on what drives attitudes towards
vigilantism, particularly in a country that has been affected by a sharp rise
in lynching episodes. Our paper should appeal especially to researchers in
Comparative Politics and in Development Studies, particularly those interested
in extralegal violence and in Latin America. We incentivize scholars to
investigate whether other forms of vigilante violence enjoy similar levels of
popular support, as well as to study which factors increase the likelihood that
citizens will actually engage in vigilante violence.

We also provide extensive supplementary materials with data, replication code,
and additional analyses for consultation at
<https://github.com/danilofreire/lynching-experiment-brazil>. The appendix
describes the survey procedures in detail and contains further information on
the qualitative evidence we gathered from respondents.

Each of the authors confirms that this manuscript has not been previously
published and is not currently under consideration by any other journal. All of
the authors have approved the contents of this paper and have agreed to
_Research & Politics_' submission policies.

Each named author has equally contributed to conducting the underlying research
and drafting this manuscript. To the best of our knowledge, the authors have no
conflict of interest, financial or otherwise.

Author names, affiliation and contact details:

* Danilo Freire: Senior Lecturer, School of Social and Political Sciences,
  University of Lincoln, <dfreire@lincoln.ac.uk>,
  <https://danilofreire.github.io>. Corresponding author.

* David Skarbek, Associate Professor, Department of Political Science, Brown
  University, <david_skarbek@brown.edu>, <http://davidskarbek.com>.

\vspace{1cm}