### 2022-09-08 title-page/title-page.Rmd
prepare manuscript submission

---
title:  Title Page
author: 
date: 
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: onehalf
papersize: a4paper
output:
  pdf_document:
    number_sections: no
    keep_tex: no
    toc: no
    template: title-page-template.latex
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# If you need to install any package while knitting the document
r <- getOption("repos")
r["CRAN"] <- "https://cran.rstudio.com/"
options(repos = r)
```

### Manuscript Title:

\noindent Vigilantism and Institutions: Understanding Attitudes toward Lynching in Brazil

### Authors:

\noindent Danilo Freire: Senior Lecturer, School of Social and Political Sciences,
  University of Lincoln, Brayford Pool, Lincoln, Lincolnshire LN6 7TS, United Kingdom, \texttt{+44 077-630-647-58}, <dfreire@lincoln.ac.uk>, <https://danilofreire.github.io>.
  Corresponding author.

\vspace{.4cm}

\noindent David Skarbek, Associate Professor, Department of Political Science, Brown
  University, 111 Thayer Street, Third Floor, Providence, RI 02912, United States of America, `+1 401-863-2825`, <david_skarbek@brown.edu>, <http://davidskarbek.com>, <https://twitter.com/davidskarbek>.